---
title: "Nobel Prize Exploration"
author: "Mohamed Hassan-El Serafi"
date: "2023-03-27"
output: html_document
---


Working with the two JSON files available through the API at nobelprize.org, ask and answer 4 interesting questions.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```



```{r}
library(tidyverse)
library(RCurl)
library(rjson)
library(data.table)
library(httr)
library(reshape2)
library(htmlwidgets)
library(kableExtra)
```



## Nobel Prize API


```{r}
prize_url <- "https://raw.githubusercontent.com/moham6839/Extra_Credit_JSON/main/Nobel_Prize.json"

prize_json_df <- jsonlite::fromJSON(prize_url) 
new_prize_df <- prize_json_df$prizes
prize_df <- as.data.frame(new_prize_df)
DT::datatable(prize_df, filter="top")
```





```{r}
new_prize_df <- unnest(prize_df)
```





```{r}
  kbl(head(new_prize_df, n = 10)) %>%
  kable_styling(latex_options="scale_down", c("striped", "hover", "condensed", full_width=F))
```




# Nobel Laureates API


```{r}
laureate_url <- "https://raw.githubusercontent.com/moham6839/Extra_Credit_JSON/main/Nobel_Laureate.json"
laureate_json_df <- jsonlite::fromJSON(laureate_url) 
new_laureate_df <- laureate_json_df$laureates
laureate_df <- as.data.frame(new_laureate_df)
DT::datatable(laureate_df, filter="top")
```

```{r}
colnames(laureate_df)
```





```{r}
new_laureate_df <- unnest(laureate_df)
new_laureate_df2 <- unnest_wider(new_laureate_df, `affiliations`, names_sep = "_")
```


```{r}
  kbl(head(new_laureate_df2, n = 10)) %>%
  kable_styling(latex_options="scale_down", c("striped", "hover", "condensed", full_width=F))
```







1. Which country has the most Nobel Laureates?


```{r}
most_laureates <- new_laureate_df2 %>%
  filter(!is.na(bornCountry)) %>%
  group_by(bornCountry) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
```


```{r}
pct_most_laureates <- most_laureates %>%
  mutate(pct_breakdown = 100 * (most_laureates$count / nrow(new_laureate_df2))) %>%
  top_n(20) %>%
  arrange(desc(pct_breakdown))
pct_most_laureates
```

```{r}
ggplot(pct_most_laureates, aes(x=reorder(bornCountry, -pct_breakdown), y=pct_breakdown)) +
  geom_bar(stat="identity", position="dodge") +
  coord_flip()
```

Close to 30% of all Nobel Laureates were born in the U.S., followed by Great Britain and Germany.


2. What is the percentage of Nobel Laureates that are still alive?

```{r}
still_alive <- new_laureate_df2 %>%
  filter(died == "0000-00-00") %>%
  group_by(died) %>%
  summarise(count=n()) %>%
  arrange(desc(died))
still_alive
```



```{r}
pct_still_alive <- (still_alive$count / nrow(new_laureate_df2)) * 100
pct_still_alive
```

Approximately 33% of Nobel Laureates are still alive.




3. Which category has the most female Nobel Laureates winners?

```{r}
category_breakdown <- new_laureate_df2 %>%
  filter(gender == "female") %>%
  group_by(category) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
category_breakdown
```





```{r}
category_breakdown %>%
ggplot(aes(x = reorder(category, -count), y=count)) +
  geom_bar(stat='identity', position='dodge', width=0.5) +
  labs(title = "Female Nobel Laureates",
       y="Number of Female Nobel Laureates",
       x="Category")
```

The category that has the most female Nobel Laureates is Peace, followed by Literature and Medicine.



4. What is the percentage of the Nobel Prize awards that had shared winners?



```{r}
shared_prize <- new_prize_df %>%
  count(share) 
shared_prize
```


```{r}
shared_prizes2 <- shared_prize %>%
  mutate(pct_shared = 100 * (shared_prize$n / nrow(new_prize_df))) %>%
  arrange(desc(pct_shared))
shared_prizes2
```




```{r}
shared_prizes2 %>%
ggplot(aes(x = reorder(share, -pct_shared), y=pct_shared)) +
  geom_bar(stat='identity', position='dodge', width=0.5) +
  labs(title = "Nobel Prizes Winners",
       y="Percentage Breakdown",
       x="Number of Times Shared")
```




```{r}
shared_breakdown <- new_prize_df %>%
  summarise(single = sum(share == 1),
            multiple = sum(share > 1))
shared_breakdown
```


```{r}
pct_shared_winners <- (shared_breakdown$multiple / sum(shared_breakdown)) * 100
pct_shared_winners
```


Approximately 64% of all Nobel Prizes were shared by at least 2 or more people. 



